---
title: "2019-03-13-pheidole-flye-LD0.3-pca-plots"
author: "EmelineFavreau"
date: "13 March 2019"
output:
  pdf_document: default
---
There are 115 samples from 3 countries. Using biallelic SNPs (from 109 pheidole, 364,197
 variants biallelic SNPs supported by 75% of samples, minor allele frequency >0.05, LD p
runed more relaxed), we explore the population structure with PCA from plink.

```{r eval = TRUE, echo = FALSE, include = FALSE}
# check the working directory
basic_libraries <- c("ggplot2", "readr")
for (lib in basic_libraries) {
  if (require(package = lib, character.only = TRUE)) {
    print("Successful")
  } else {
    print("Installing")
    install.packages(lib)
    library(lib, character.only = TRUE )
  }
}
```

Looking through all PC axes up to 20, there is no split between monogynous and polygynous.

```{r eval = TRUE, echo = FALSE}
# import the pca file
pheidole.eigenvec <- read.table("result/pheidole-0.3-PCA.eigenvec", header = TRUE)
# update social forms
pheidole.eigenvec$FID <- as.character(pheidole.eigenvec$FID)
pheidole.eigenvec$FID[pheidole.eigenvec$FID == "A56-N"] <- "A56-M"
pheidole.eigenvec$FID[pheidole.eigenvec$FID == "I27-N"] <- "I27-P"
pheidole.eigenvec$FID[pheidole.eigenvec$FID == "andrea-N"] <- "andrea-P"
pheidole.eigenvec$FID[pheidole.eigenvec$FID == "muna-N"] <- "muna-P"

# import population and gyny info
pop <- read.csv("S2_pheidole_pop_paper.csv", header = TRUE)
#head(pop)
pop$sample.soc <- as.character(pop$sample.soc)
colnames(pop) <- c("FID", "population", "gyny")

# merge the two dataframes
my_df <- merge(x = pheidole.eigenvec, y = pop, intersect(names(pheidole.eigenvec), names
(pop)))
my_df$population <- as.character(my_df$population)

# Investigate all 109 samples
ggplot(my_df, aes(x = PC1, y = PC2, color = population, shape = gyny)) + 
  geom_point(size = 6, alpha = 0.6) + geom_text(aes(label = FID), hjust = 0, vjust = 0) 
+ theme_classic() + ggtitle("from non-LD variants based on 109 samples \n maf>0.05")

```



Investigating the proportion of variance explained by each PC axis. 20 axes explain 45% of the whole variance. PC1 to PC5 explain 20% of the whole variation. 

```{r eval = TRUE, echo = FALSE}
# import the pca file: top eigenvalues are written to plink.eigenval.
pheidole.eigenval <- read.table("result/pheidole-0.3-PCA.eigenval", header = FALSE)
rownames(pheidole.eigenval) <- unique(rep(paste("PC", 1:20, sep = ""), 20))
colnames(pheidole.eigenval) <- c("% of variance explained")
print(pheidole.eigenval)
sum(pheidole.eigenval$`% of variance explained`[1:5])
sum(pheidole.eigenval$`% of variance explained`)
# PC1 to PC5 explain 21% of the whole variation.
pcs_to_keep <- rownames(pheidole.eigenval)[1:2]
pheidole.eigenvec.tops <- pheidole.eigenvec[, c(colnames(pheidole.eigenvec)[1:2], pcs_to_keep)]
write.table(pheidole.eigenvec.tops, file = "result/pheidole.0.3.eigenvec.tops.csv", quote = FALSE, row.names = FALSE)
```

